Nested Stochastic Block Models Applied to the Analysis of Single Cell Data

### 2021-04-26 manuscript.tex
more methods

% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{authblk}
\usepackage[numbers]{natbib}
\usepackage{xr}
\externaldocument{supplementary}

@@

\subsection*{Analysis of Visium H\&E data}
\textcolor{red}{TOBEDONE}

@@

\subsection*{Label transfer}
\textcolor{red}{TO BE DONE}

\subsection*{Analysis of cluster consistency}

Count matrices were downloaded from GEO using the following accession numbers: GSE133535 (Chromium 10Xv3), GSE133543 (Quartz-seq2), GSE133542 (MARS-seq) and GSE133541 (iCELL8). Data were processed according to the methods in the original paper \cite{mereu_2020}. Briefly, cells with less than 10,000 total number of reads as well as the cells having less than 65\% of the reads mapped to their reference genome were discarded. Cells in the 95th percentile of the number of genes/cell and those having less than 25\% mitochondrial gene content were included in the downstream analyses. Genes that were expressed in less than five cells were removed. Data were normalized and log-transformed, highly variable genes were detected at minimal dispersion equal to 0.5. Neighbourhood graph was built using 30 principal components and 20 neighbours. nSBM was completed with 3 initialisations.

>>>>

% !TEX encoding = UTF-8 Unicode

\documentclass[10pt]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{authblk}
\usepackage[numbers]{natbib}
\usepackage{xr}
\externaldocument{supplementary}

@@

\subsection*{Analysis of Visium H\&E data}
Data were retrieved using \texttt{squidpy.datasets.visium\_hne\_adata()} built-in function, without further processing. Leiden clustering was performed using \texttt{schist.inference.leiden()} function, allowing for 100 initializations, with resolutions $\gamma = 1$ and $\gamma = 0.5$.

@@

\subsection*{Processing of PBMC data from various platforms}
Count matrices were downloaded from GEO using the following accession numbers: GSE133535 (Chromium 10Xv3), GSE133543 (Quartz-seq2), GSE133542 (MARS-seq) and GSE133541 (iCELL8). Data were processed according to the methods in the original paper \cite{mereu_2020}. Briefly, cells with less than 10,000 total number of reads as well as the cells having less than 65\% of the reads mapped to their reference genome were discarded. Cells in the 95th percentile of the number of genes/cell and those having less than 25\% mitochondrial gene content were included in the downstream analyses. Genes that were expressed in less than five cells were removed. Data were normalized and log-transformed, highly variable genes were detected at minimal dispersion equal to 0.5. Neighbourhood graph was built using 30 principal components and 20 neighbours.

\subsection*{Label transfer}

Processed data for MARS-seq or Quart-seq2 platforms were merged to data for 10X V3. Merged data were then processed using Harmony \cite{Korsunsky_2019} by the \texttt{scanpy.external.pp.harmony\_integrate()} function with default parameters. Cells not belonging to the 10X data were assigned an "Unknown" label. We calculated cell affinity to each annotation label using \texttt{schist.tl.calculated\_affinity()} function. We assigned the most affine annotation only to "Unknown" cells. For \emph{k}NN-based procedure, we built a \emph{k}NN graph on the merged data using \texttt{pynndescent } library on the 10XV3 subset of cells in the merged data, then we assigned "Unknown" cells to the closest entry in the graph. Assignment by \texttt{scanpy.tl.ingest()} was performed using default parameters.
